Internal Gravity Wave Breaking in White Dwarf Binaries

### 2019-08-08 writeups/Su_IGW_break.tex
bleh

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

In the small perturbation limit, where flow velocities are small compared to the
characteristic space and time scales $\pd{}{t} \gg \bm{u}' \cdot
\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
linear regime up to undetermined amplitude $A$ is given by
\citep{drazin,sutherland0}:
\begin{equation}
    u'_z\p{x, z, t} = Ae^{z/2H}e^{i\p{k_{x}x + k_{z}z - \omega t}},
        \label{eq:k0z_sign}
\end{equation}
where $\omega$ satisfies dispersion relation
\begin{equation}
    \omega^2 = \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
        \label{eq:disp_rel}
\end{equation}
We take
\begin{equation}
    N^2 \equiv g^2\p{\rd{\rho}{P} - \frac{1}{c_s^2}} = \frac{g}{H},
\end{equation}
the \emph{Brunt-V\"ais\"al\"a frequency} to be constant in the incompressible
limit $c_s \to \infty$ where $c_s$ is the sound speed. Other dynamical
quantities are simply related to $u'_z$ (see e.g.\ \cite{sutherland0}).

@@

    \item The time-averaged total $x$-momentum flux $F$ carried in the $\bm{z}$
        direction can be computed. Since the linear solution consists of terms
        separable as $f(x, z, t) = f(z)e^{i(k_{x}x - \omega t)}$, $x$
        averaging and time averaging are equivalent. We may then write
        \begin{align}
            F &\equiv \ev{\rho u_{x}' u_{z}'}_x \equiv
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x,
                    \label{eq:F_def},\\
                &\approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
                    \label{eq:S_lin}
        \end{align}
        where $\ev{\dots}_x$ denotes $x$-averaging. Note that indeed for an
        upwards propagating IGW $k_z < 0, k_x > 0$, we have $F > 0$.
\end{itemize}

@@

Since we are only concerned with the behavior of the wave at high $z$ where wave
breaking occurs, we generate waves in the most numerically convenient way. As
long as physically represesentative IGW are generated and wave dynamics are only
examined far from the forcing region, the exact excitation mechanism is not
important for our conclusions.

@@

As not to interfere with the incompressibility constraint, we force the system
on the density equation. We implement forcing with strength $C$ localized around
$z_0$ with width $\sigma$ by replacing \autoref{eq:nl_up_density} with
\begin{equation}
    \md{\Upsilon}{t} + u_{z}'\pd{\ln \overline{\rho}}{z}
        = Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
            \cos \p{k_{x}x - \omega t}.\label{eq:vol_drive}
\end{equation}
Using a narrow Gaussian profile excites a broad $z$ power spectrum, but only the
$k_{z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the given
$k_{x}, \omega(k_{x}, k_{z})$ will propagate.

In the linearized system, the effect of this forcing can be solved analytically
up to good accuracy: we first approximate the driving term using
$e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
z_0)$, the $\sigma \to 0$ limit\footnote{In practice, $\sigma$ must be large
enough to be numerically resolved by the spectral code.}. This system is solved
exactly by matching the two plane wave solutions, above and below $z_0$. We may
then approximately relax the solution to nonzero $\sigma$: an extra factor of
$e^{-\frac{(k_{z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
solution (evaluating the Fourier Transform of $e^{-\frac{(z -
z_0)^2}{2\sigma^2}}$ at $k_z$), and we obtain
\begin{align}
    u_{z}'&(x, z, t) ={} \frac{Cgk_{x}^2}{\omega^2}
        \frac{1}{2ik_{z}}\frac{e^{\frac{-(k_{z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}} \times\nonumber\\
        &{}\begin{cases}
        e^{\frac{z - z_0}{2H}}e^{i\p{k_{x}x + k_{z}(z - z_0) - \omega t
            + \frac{1}{2k_{z}H}}}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}e^{i\p{k_{x}x - k_{z}(z - z_0) - \omega t
            + \frac{1}{2k_{z}H}}}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}
The $z > z_0$ region models an upwards-propagating IGW wavetrain excited deep
in the atmosphere.

\subsection{Wave Breaking Location}\label{ss:wave_breaking}

@@

This is in agreement with the breaking location predicted by the wave-induced
mean flow, given by \citep{eliassen_palm_cite,sutherland0}:
\begin{equation}
     \ev{u_{x}} \equiv \overline{U} = \frac{1}{L_x}
        \int\limits u_{x}\;\mathrm{d}x = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}},
        \label{eq:mean_flow}
\end{equation}
This formula is analogous to Stokes' drift for surface waves and can be derived
by considering the propagation of $F$ (\autoref{eq:F_def}) into a medium at
rest. Solving for $\overline{U} = \omega / k_x$ the critical layer condition
(\autoref{eq:u_c} and discussion in \autoref{ss:crit_layer}), we obtain
$\abs{\frac{u_{x}'k_{x}}{\omega}} = \abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$,
equal to the nonlinearity criterion \autoref{eq:nl}. This condition predicts
where the wave's self-acceleration alone produces a critical layer.

@@

% TODO physical vs simulation domain
Denote $L_x, L_z$ to be the physical dimensions of the simulation domain. Our
simulation domain is smaller than our full physical domain, so we adopt
reflection-suppressing boundary conditions. We use periodic boundary conditions
in the $x$ direction and damping layers (described in \autoref{ss:damping}) in
the $z$ direction to damp perturbations exiting the top and bottom of the
domain. We use a Fourier basis in both the $x$ and $z$ direction. We varied the
number of $x, z$ modes (denoted $N_x, N_z$ respectively), and used $3/2$
dealiasing \citep{boyd}.

The geometry of our simulation domain is fixed by three parameters: $L_x, L_z$,
and $z_0$ the forcing location. We use $z_0 = 0.2L_z$ to force sufficiently far
from the lower damping zone and permit sufficient room for the upwards-moving
wave to grow at $\propto e^{z/2H}$. We choose $L_z = 10H$ to give $\sim e^3$
apmlitude growth between the damping zones. Finally, we want similar grid
spacing $\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$. For computational savings, we
fix $N_z = 4N_x$. We use $L_x = 4H = \frac{2}{5}L_z$.

@@

    \item $\omega$: We choose $\omega$ by evaluating dispersion relation
        $\omega\p{k_{x}, k_{z}}$ for some desired $k_{z}$.

        To choose $k_z$, we note astrophysical IGWs generally satisfy $\omega
        \ll N$, or equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since
        $H \lesssim R$, we aim to study $k_{z}H \gg 1$. However, to satisfy this
        well is computationally expensive: $L_z \gg H$ is required to give waves
        ample room to grow within the simulation domain, but we want $k_{z}$ to
        be sufficiently separated from the grid spacing $\sim L_z / N_z$ that
        nonlinear effects can be seen in later simulations (see
        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{z}H
        \gg 1$; we choose $k_{z}H = 1$.

    \item $C$: We first choose $C$ forcing strength such that $\xi_z k_z \ll 1$
        is satisfied everywhere throughout the domain. This constrains $C$ by
        \autoref{eq:uz_lin}.

@@

To imitate an infinite fluid using a finite simulation domain, we use periodic
boundary conditions in the $x$ direction and damping layers at the top and
bottom of the $z$ direction. This damps waves that reach the edge of the
simulation domain without inducing nonphysical reflection. Specifically, we
replace material derivatives in \autoref{se:nl_upsilon} with:
\begin{align}
    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\nonumber\\
    \Gamma(z) &= \frac{1}{2\tau}\s{2,+ \tanh \frac{z - z_T}{\Delta z}
        + \tanh \frac{z_B - z}{\Delta z}},\label{eq:Gamma}
\end{align}
where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zone.
This strongly damps perturbations below $z_B$ and above $z_T$ with damping time
$\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
governed by $\Delta z$; we use $\Delta z = 0.025L_z$. This prescription, used in
similar studies \citep{lecoanet_damp}, has the advantage of being smooth,
important for spectral methods. Further details of our implementation of the
fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.

@@

We describe the results of a simulation satisfying the parameter choices in
\autoref{ss:numerics} such that $\xi_z k_z \ll 1$ everywhere.

During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be excited.
The incompressibility constraint further yields a complete analytical solution
for flow velocities $\bm{u}_{al}'(x, z, t)$. The amplitude of the observed IGW
in the simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}'$
over some region $z \in [z_b, z_t]$ can be estimated with estimator
$\hat{A}_i(t)$ (subscript $i$ denotes incident wave)
\begin{equation}
    \hat{A}_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
        \overline{\rho}\p{\bm{u} \cdot \bm{u}_{al}'}\;\mathrm{d}x\mathrm{d}z}
        {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
        \overline{\rho}(u_{al}')^2\;\mathrm{d}x\mathrm{d}z}
        \label{eq:ahat_def}
\end{equation}
If $\bm{u} = \bm{u}_{al}'$, then $\hat{A}_i(t) = 1$. The energy norm is a
traditional normalization choice such that the overlap between $\bm{u},
\bm{u}_{al}' \propto e^{z/2H}$ is evenly weighted throughout the integration
region.

@@

The analytical theory also predicts that the horizontal momentmum flux $F(z, t)$
should be spatially flat between the forcing zone where it is generated and the
damping zone where it is dissipated (\autoref{eq:F_def}). We make may a slightly
more accurate estimate of $F$ the flux carried in the IGW by directly
substituting $\bm{u}_{al}'$ into \autoref{eq:F_def}. Calling this estimate
\begin{equation}
    F_{al}' \equiv \frac{1}{L_x}
        \int\limits_{0}^{L_x} \overline{\rho} u_{al,x}'u_{al,z}'\;\mathrm{d}x,
\end{equation}
we may then measure the agreement of our simulation with analytical expectation
by computing
\begin{equation}
    \hat{F}(z, t) \equiv \frac{
        \frac{1}{L_x}\int\limits_0^{L_x}\rho u_xu_z\;\mathrm{d}x}
        {F_{al}'}.\label{eq:sbm_def}
\end{equation}
For our weak forcing simulation, $\hat{F} = 1$ is expected between $z_0, z_T$,
and indeed we observe agreement with this in \autoref{fig:lin_fluxes}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
    \caption{\autoref{eq:sbm_def} as a function of $z$ at select times $t$. As
    initial transients die out, $\hat{F}(z, t) = 1$ to very good agreement above
    the forcing zone $z > z_0 = 2H$ and below the damping zone $z \lesssim z_T =
    9.5H$. The flux excited in the forcing zone is transported without loss to
    the top of the domain, where it is dissipated by the damping layer (see
    \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
\end{figure}

@@

The evolution of $z_c$ depends on the deposited flux $\Delta F(t)$ per
\autoref{eq:zc_anal}. To estimate $\Delta F(t)$ across the critical layer $z_c$
from the data, we defined:
\begin{align}
    F_{<}(t) &= \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
        ,\label{eq:sbelow_def}\\
    F_>(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
        \label{eq:sabove_def},\\
    \Delta F(t) &\equiv F_>(t) - F_{<}(t).\label{eq:ds_def}
\end{align}
$\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
Below the critical layer, we average over an interval of length $H =
\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
necessary to make a measurement of the incident flux unaffected by the
turbulence within the critical layer. The height of the critical layer is
limited by $\mathrm{Ri} \lesssim 1$, which bounds its vertical extent $\sim
\frac{1}{k_{z}}$. Here, we found $\Delta z = \frac{3}{k_z}$ was necessary to be
sufficiently offset from strong fluctuations near the critical layer.

Above the critical layer, waves decay over a much smaller vertical extent as
they have larger $z$ wavenumbers thanks to the shear flow, so we average over a
$\Delta z$. Accurate measurement of these fluctuations above the critical layer
is important to measuring the transmission feature accurately, seen in
\autoref{fig:nl_fluxes} to be weak and attenuate quickly.

Finally, we may plot the observed $z_c$ from data against two simple predictors:
(i) integration of \autoref{eq:zc_anal} using the measured $\Delta F(t)$, and
(ii) substituting the time-averaged $\ev{\Delta F}_t$ into \autoref{eq:zc_sol}.
Since $z_c(t)$ is both stable and less well-defined at early times (when the
critical layer is thick and transient behavior is still strong), we instead
integrate backwards from the end of the simulation, using $z_c(t_f)$ as initial
condition. The resulting predictors are depicted in \autoref{fig:nl_front}. The
good agreement between the evolution of $z_c(t)$ and its estimate via $\Delta
F(t)$ and \autoref{eq:zc_anal} are noteworthy. Further of interest is the
discrepancy between the time-averaged predictor and the data. The general
agreement clearly demonstrates $\Delta F < F_{al}'$ and that incomplete
absorption is observed. Moreover, its significant underprediction of critical
layer motion at early times contrasts with its comparatively satisfactory
prediction at later times to indicate that $\Delta F(t)$ must vary significantly
over time. Both of these observations indicate a careful characterization of
reflectivity is necessary.

@@

We identify two manifestations of non-absorptive behavior: (i), the presence of
a reflected wave with wave vector $\bm{k} = k_{x}\bm{x} - k_{z}\bm{z}$, and (ii)
the amount of horizontal momentum flux $F$ reflected off the critical layer.
These two quantities need not agree exactly if some reflected flux is in
higher-order modes, which is observed. Both are of physical interest, however:
he reflected wave amplitude may be of physical interest in setting up standing
modes in a realistic star, while the reflected flux is important for accurately
tracking angular momentum transfer during synchronization.

@@

Since reflectivity depends sensitively on accurate measurements of $\hat{A}_i,
\hat{A}_r$, we remark \autoref{eq:ahat_def} ensures orthogonality between
$k_{x}\bm{x} \pm k_{z}\bm{z}$ modes. The integral for $\hat{A}_r$ is also
performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat over time, we perform time averaging
over interval approximately $8\pi/\omega$, denoted by angle brackets. We can
then define the amplitude
reflectivity
\begin{equation}
    \mathcal{R}_A(t) \equiv \frac{\ev{\hat{A}_r}(t)}{\ev{\hat{A}_i}(t)}
        .\label{eq:Ra_def}
\end{equation}
% 7/N per timestep, 13 timestep smoothing kernel, 24/N period

>>>>

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\left\langle#1\right\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

In the small perturbation limit, where flow velocities are small compared to the
characteristic space and time scales $\pd{}{t} \gg \bm{u}' \cdot
\bm{\nabla}$, we may linearize \autoref{se:nl_upsilon}. The solution in this
linear regime up to undetermined amplitude $A$ is given by
\citep{drazin,sutherland0}:
\begin{equation}
    u'_z\p{x, z, t} = Ae^{z/2H}\cos\p{\p{k_{x}x + k_{z}z - \omega t}},
        \label{eq:k0z_sign}
\end{equation}
where $\omega$ satisfies dispersion relation
\begin{equation}
    \omega^2 = \frac{N^2k_{x}^2}{k_{x}^2 + k_{z}^2 + \frac{1}{4H^2}}.
        \label{eq:disp_rel}
\end{equation}
We take
\begin{equation}
    N^2 \equiv g^2\p{\rd{\rho}{P} - \frac{1}{c_s^2}} = \frac{g}{H},
\end{equation}
the \emph{Brunt-V\"ais\"al\"a frequency} to be constant in the incompressible
limit $c_s \to \infty$ where $c_s$ is the sound speed. Other dynamical
quantities are simply related to $u'_z$ (see e.g.\ \cite{sutherland0}).

@@

    \item The time-averaged total $x$-momentum flux $F$ carried in the $\bm{z}$
        direction can be computed. Since the linear solution consists of terms
        separable as $f(x, z, t) = f(z)\cos\p{k_{x}x - \omega t}$, $x$
        averaging and time averaging are equivalent. $x$ averaging is preferred
        both since periodic boundary conditions pick a natural averaging length
        and since it does not introduce extra temporal smoothing. We may then
        write
        \begin{align}
            F &\equiv \ev{\rho u_{x}' u_{z}'}_x \equiv
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_{x}'u_{z}'\;\mathrm{d}x,
                    \label{eq:F_def},\\
                &\approx -\frac{A^2}{2}\overline{\rho}_0\frac{k_{z}}{k_{x}},
                    \label{eq:S_lin}
        \end{align}
        where $\ev{\dots}_x$ denotes $x$-averaging. Note that indeed for an
        upwards propagating IGW $k_z < 0, k_x > 0$, we have $F > 0$.
\end{itemize}

@@

Since we are only concerned with the behavior of the wave at high $z$ where wave
breaking occurs, we generate waves at low $z$ in the most numerically convenient
way. As long as physically represesentative, upwards-propgaating IGW are
generated and wave dynamics are only examined far from the forcing region, the
exact excitation mechanism is not important for our conclusions.

@@

As not to interfere with the incompressibility constraint, we force the system
on the density equation. We implement forcing with strength $C$ localized around
$z_0$ with small width $\sigma$ by replacing \autoref{eq:nl_up_density} with
\begin{equation}
    \md{\Upsilon}{t} + u_{z}'\pd{\ln \overline{\rho}}{z}
        = Ce^{-\frac{(z - z_0)^2}{2\sigma^2}}
            \cos \p{k_{x}x - \omega t}.\label{eq:vol_drive}
\end{equation}
Using a narrow Gaussian profile excites a broad $z$ power spectrum, but only the
$k_{z}$ satisfying dispersion relation \autoref{eq:disp_rel} for the given
$k_{x}$, $\omega(k_{x}, k_{z})$ will propagate.

In the linearized system, the effect of this forcing can be solved analytically
up to good accuracy: we first approximate the driving term using
$e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
z_0)$, the $\sigma \to 0$ limit\footnote{In practice, $\sigma$ must be large
enough to be numerically resolved by the spectral code.}. This system is solved
exactly by matching the two plane wave solutions, above and below $z_0$. We may
then approximately relax the solution to nonzero $\sigma$: an extra factor of
$e^{-\frac{(k_{z}\sigma)^2}{2}}$ arises compared to the $\delta$-function
solution (evaluating the Fourier Transform of $e^{-\frac{(z -
z_0)^2}{2\sigma^2}}$ at $k_z$), and we obtain
\begin{align}
    u_{z}'&(x, z, t) ={} \frac{Cgk_{x}^2}{\omega^2}
        \frac{1}{2ik_{z}}\frac{e^{\frac{-(k_{z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}} \times\nonumber\\
        &{}\begin{cases}
        e^{\frac{z - z_0}{2H}}\cos\p{k_{x}x + k_{z}(z - z_0) - \omega t
            + \frac{1}{2k_{z}H}}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}\cos\p{k_{x}x - k_{z}(z - z_0) - \omega t
            + \frac{1}{2k_{z}H}}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}
The $z > z_0$ region models an upwards-propagating IGW wavetrain excited deep
in the atmosphere.

\subsection{Wave Breaking Height}\label{ss:wave_breaking}

@@

This is in agreement with the breaking location predicted by the wave-induced
mean flow, given by \citep{eliassen_palm_cite,sutherland0}:
\begin{equation}
     \ev{u_{x}} \equiv \overline{U} = \frac{1}{L_x}
        \int\limits u_{x}\;\mathrm{d}x = \frac{\ev{u_{x}u_{z}}_x}{c_{g,z}},
        \label{eq:mean_flow}
\end{equation}
This formula is analogous to Stokes' drift for surface waves and can be derived
by considering the propagation of $F$ (\autoref{eq:F_def}) into a medium at
rest. Solving for where $\overline{U} = \omega / k_x$ the critical layer
condition is fulfilled (\autoref{eq:u_c} and discussion in
\autoref{ss:crit_layer}), we obtain $\abs{\frac{u_{x}'k_{x}}{\omega}} =
\abs{\xi_x k_{x}} = \abs{\xi_z k_z} \sim 1$, equal to the nonlinearity criterion
\autoref{eq:nl}. This condition predicts where the wave's self-acceleration
alone produces a critical layer.

@@

In a realistic WD, IGW are excited very deep in the WD and propagate a
substantial number of density scale heights before reaching the breaking zone.
This propagation is very nearly exactly linear until just a few scale heights
below the breaking region. In the interest of computational feasibility, we
restrct the simulation domain to the upper WD fluid, of much smaller vertical
extent than the full physical domain. To further decrease computational cost, we
restrict ourselves to 2D, plane-parallel IGW rather than full 3D stellar
g-modes. While it is well-known wave breaking is a 3D process
(\cite{klostermeyer,winters1994}), the dynamical effect of the breaking process
is likely to be similar in 2D.

Denote $L_x, L_z$ to be the dimensions of the simulation domain. As the physical
domain extends beyond $L_x, L_z$, we must take care to suppress reflections. We
use periodic boundary conditions in the $x$ direction and damping layers
(described in \autoref{ss:damping}) in the $z$ direction to damp perturbations
incident on the top and bottom of the simulation. We use a Fourier basis in both
the $x$ and $z$ direction. We varied the number of $x, z$ modes (denoted $N_x,
N_z$ respectively), and used $3/2$ dealiasing \citep{boyd}.

The geometry of our simulation domain is fixed by one further parameter: $z_0$
the forcing location. We use $z_0 = 0.2L_z$ to force sufficiently far from the
lower damping zone and permit sufficient room for the upwards-moving wave to
grow at $\propto e^{z/2H}$. We choose $L_z = 10H$ to give $\sim e^3$ apmlitude
growth between the damping zones. Finally, we want similar grid spacing
$\frac{L_x}{N_x} \sim \frac{L_z}{N_z}$, guided by the intuition that turbulence
is approximately isotropic. For computational savings, we fix $L_x = 4H < L_z$,
a compromise between using fewer $N_x$ modes and using as large of an $L_x$ as
possible (in stars, $L_x \sim 2\pi R$ the circumference). We then fix ratio $N_z
/ N_x = 4$.

@@

    \item $\omega$: We choose $\omega$ by evaluating dispersion relation
        $\omega\p{k_{x}, k_{z}}$ for a desired $k_{z}$.

        To choose $k_z$, we note astrophysical IGWs generally satisfy $\omega
        \ll N$, or equivalently $\frac{k_r}{k_{\perp}} \sim k_rR \gg 1$. Since
        $H \lesssim R$, we aim to study $k_{z}H \gg 1$. However, to satisfy this
        well requires high resolution simulations: $L_z \gg H$ is required to
        give waves ample room to grow within the simulation domain, but we want
        $k_{z}$ to be sufficiently separated from the grid spacing $\sim L_z /
        N_z$ that nonlinear effects can be seen in later simulations (see
        \autoref{ss:nl_ns}). Thus, $N_z$ sets how well we can satisfy $k_{z}H
        \gg 1$; we choose $k_{z}H = 1$ as a compromise.

    \item $C$: In our linear simulations, we first choose $C$ forcing strength
        such that $\xi_z k_z \ll 1$ is satisfied everywhere in the simulation
        domain. This constrains $C$ by \autoref{eq:uz_lin}.

@@

We aim to damp waves that reach the edge of the simulation domain without
inducing nonphysical reflection. To do so, we replace material derivatives
in \autoref{se:nl_upsilon} with:
\begin{align}
    \md{}{t} &\to \md{}{t} + \Gamma(z)\Upsilon,\nonumber\\
    \Gamma(z) &= \frac{1}{2\tau}\s{2,+ \tanh \frac{z - z_T}{\Delta z}
        + \tanh \frac{z_B - z}{\Delta z}},\label{eq:Gamma}
\end{align}
where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zone.
This strongly damps perturbations below $z_B$ and above $z_T$ with damping time
$\tau$, negligibly affects dynamics between $z_B, z_T$ and has transition width
governed by $\Delta z$; we use $\Delta z = 0.025L_z$. This prescription, used in
similar studies \citep{lecoanet_damp}, has the advantage of being smooth,
important for spectral methods. Further details of our implementation of the
fluid equations in Dedalus are described in Appendix~\ref{se:strat_impl}.

@@

We describe the results of a simulation satisfying the parameter choices in
\autoref{ss:numerics} such that $\xi_z k_z \ll 1$ everywhere in the simulation
domain.

During the simulation, we expect IGW of form \autoref{eq:uz_lin} to be excited.
Invoking incompressibility, we obtain a complete analytical solution for flow
velocities $\bm{u}_{al}'(x, z, t)$. The amplitude of the observed IGW in the
simulation field $\bm{u}$ relative to analytical solution $\bm{u}_{al}'$ over
some region $z \in [z_b, z_t]$ can be estimated with estimator $\hat{A}_i(t)$
(subscript $i$ denotes incident wave)
\begin{equation}
    \hat{A}_i(t) = \frac{\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
        \overline{\rho}\p{\bm{u} \cdot \bm{u}_{al}'}\;\mathrm{d}x\mathrm{d}z}
        {\int\limits_{z_b}^{z_t}\int\limits_0^{L_x}
        \overline{\rho}\p{u_{al}'}^2\;\mathrm{d}x\mathrm{d}z}
        \label{eq:ahat_def}
\end{equation}
If $\bm{u} = \bm{u}_{al}'$, then $\hat{A}_i(t) = 1$. The energy norm is a
traditional normalization choice such that the overlap between $\bm{u},
\bm{u}_{al}' \propto e^{z/2H}$ is evenly weighted throughout the integration
region.

@@

The analytical theory also predicts that the horizontal momentmum flux $F(z, t)$
should be spatially flat between the forcing zone where it is generated and the
damping zone where it is dissipated (\autoref{eq:F_def}). We make may a slightly
more accurate estimate of the flux carried in the IGW by directly substituting
$\bm{u}_{al}'$ into \autoref{eq:F_def}. Calling this estimate
\begin{equation}
    F_{al}'(z) \equiv \ev{\overline{\rho} u_{al,x}'u_{al,z}'}_x
\end{equation}
we may then measure the agreement of our simulation with analytical expectation
by computing
\begin{equation}
    \hat{F}(z, t) \equiv \frac{\ev{\rho u_xu_z}_x}
        {F_{al}'}.\label{eq:sbm_def}
\end{equation}
For our weak forcing simulation, $\hat{F}(z) = 1$ is expected between $z_0, z_T$
after initial transients dissipate, and indeed we observe agreement with this in
\autoref{fig:lin_fluxes}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{plots/lin_fluxes.png}
    \caption{\autoref{eq:sbm_def} as a function of $z$ at select times $t$. As
    initial transients die out, $\hat{F}(z, t) = 1$ to very good agreement above
    the forcing zone $z > z_0 = 2H$ and below the damping zone $z \lesssim z_T =
    9.5H$. The flux excited in the forcing zone is transported without loss to
    the top of the domain, where it is dissipated by the damping layer (see
    \autoref{ss:damping}) without reflection.}\label{fig:lin_fluxes}
\end{figure}

@@

The evolution of $z_c$ depends on the deposited flux $\Delta F(t)$ per
\autoref{eq:zc_anal}. To estimate $\Delta F(t)$ across the critical layer $z_c$
from the data, we defined:
\begin{align}
    F_{<}(t) &= \ev{F(z)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
        ,\label{eq:sbelow_def}\\
    F_>(t) &= \ev{\z{F(z): F(z) < 0}}_{z \in [z_c, z_c + \Delta z]},
        \label{eq:sabove_def},\\
    \Delta F(t) &\equiv F_>(t) - F_{<}(t).\label{eq:ds_def}
\end{align}
$\ev{\dots}_{z \in [z_a, z_b]}$ denotes averaging over interval $[z_a, z_b]$.
Below the critical layer, we average over an interval of length
$\frac{2\pi}{k_{z}}$ a full vertical wavelength. The offset $\Delta z$ is
necessary to make a measurement of the incident flux unaffected by the
turbulence within the critical layer itself. The height of the critical layer is
limited by $\mathrm{Ri} \lesssim 1$, which bounds its vertical extent $\sim
\frac{1}{k_{z}}$. We empirically found an offset of $\Delta z = \frac{3}{k_z}$
was necessary to be sufficiently far from strong fluctuations near the critical
layer.

Above the critical layer, waves decay over a much smaller vertical extent as
they have larger $z$ wavenumbers thanks to the shear flow Doppler shift.
Accurate measurement of these fluctuations above the critical layer is important
to measuring the transmission feature accurately, seen in
\autoref{fig:nl_fluxes} to be weak and attenuate quickly.

Finally, we may plot the measured $z_c$ against two semi-analytic predictors:
(i) integration of \autoref{eq:zc_anal} using the measured $\Delta F(t)$, and
(ii) substituting the time-averaged $\ev{\Delta F}_t$ (over the entire length of
the simulation) into \autoref{eq:zc_sol}. Since $z_c(t)$ is both stable and less
well-defined at early times (when the critical layer is thick and transient
behavior is still strong), we instead integrate backwards from the end of the
simulation, using $z_c(t_f)$ as initial condition. The resulting predictors are
depicted in \autoref{fig:nl_front}. The good agreement between the evolution of
$z_c(t)$ and its estimate via $\Delta F(t)$ and \autoref{eq:zc_anal} are
noteworthy.

Also of interest is the behavior of the time-averaged $\ev{\Delta F}_t =
0.71F_{al}'$ predictor, which serves mostly as a fiducial in the plot. The
general agreement clearly demonstrates $\Delta F < F_{al}'$ and that momentum
flux absorption is incomplete. However, its underprediction of critical layer
velocity at early times contrasts with its overprediction at later times. This
suggests that momentum flux absorption changes significantly over time, which is
in line with the observations in the subsequent \autoref{ss:reflectivity}.

@@

We identify two manifestations of non-absorptive behavior: (i), the presence of
a reflected wave with wave vector $\bm{k} = k_{x}\bm{x} - k_{z}\bm{z}$, and (ii)
the amount of horizontal momentum flux $F$ reflected/transmitted by the critical
layer. The reflected wave amplitude and reflected flux need not agree exactly if
some reflected flux is in higher-order modes, which is indeed the case in our
simulations. Both are of physical interest, however: the reflected wave
amplitude may be of physical interest in setting up standing modes in a
realistic star, while the flux transfer properties is important for accurately
tracking angular momentum transfer during synchronization.

@@

Since reflectivity depends sensitively on accurate measurements of $\hat{A}_i,
\hat{A}_r$, we remark \autoref{eq:ahat_def} ensures orthogonality between
$k_{x}\bm{x} \pm k_{z}\bm{z}$ modes. The integral for $\hat{A}_r$ is also
performed over $z \in [z_0 + 3\sigma, z_0 + 3\sigma + H]$. Since
$\hat{A}_i(t), \hat{A}_r(t)$ vary somewhat strongly over time, we perform time
averaging over interval approximately $8\pi/\omega$, denoted by angle brackets.
We can then define the amplitude reflectivity
\begin{equation}
    \mathcal{R}_A(t) \equiv \frac{\ev{\hat{A}_r}(t)}{\ev{\hat{A}_i}(t)}
        .\label{eq:Ra_def}
\end{equation}
% 7/N per timestep, 13 timestep smoothing kernel, 24/N period
